GMSE: an R package for generalised management strategy evaluation

### 2018-05-01 notebook/ms/SI5.Rmd
Update author list

---
title: 'GMSE: an R package for generalised management strategy evaluation'
author: A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;,
  Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, O. Sarobidy Rakotonarivo&#xb9;,
  Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes: null
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
subtitle: Supporting Information 5
biblio-style: apalike
---

@@

Here we closely follow [a tutorial from the FLR project](http://www.flr-project.org/doc/Modelling_stock_recruitment_with_FLSR.html). To build the stock-recruitment model, the following packages are needed.

```{r, eval = FALSE}
install.packages(c("ggplot2"));
install.packages(c("FLCore"), repos="http://flr-project.org/R");
install.packages(c("ggplotFL"), repos="http://flr-project.org/R");
```

To start, we need to read in the `FLCore`, `FLFleet` and `ggplotFL` libraries.

```{r}
library(FLCore);
#library(FLFleet)
#library(ggplotFL);
```

For a simplified example in GMSE, we will simulate the process of stock recruitment over multiple time steps using an example stock-recruitment model. The stock-recruitment model describes the relationship between stock-recruitment and spawning stock biomass. The sample that we will work from is a recreation of the North Sea Herring (`nsher`) dataset available in the `FLCore` package. This data set includes recruitment and spawning stock biomass data between 1960 and 2004. First, we initialise an empty `FLSR` object and read in the recreated CSV data files.

```{r}
newFL        <- FLSR(); # Initialises the empty FLSR object
rec.n        <- read.csv("data/SI5/nsher_rec.csv");
ssb.n        <- read.csv("data/SI5/nsher_ssb.csv");
print(rec.n);
```

@@

```{r}
Frec.m       <- FLQuant(rec.m, dimnames=list(age=1, year = 1960:2004));
Fssb.m       <- FLQuant(ssb.m, dimnames=list(age=1, year = 1960:2004));
Frec.m@units <- "10^3";
Fssb.m@units <- "t*10^3";

newFL@rec    <- Frec.m;
newFL@ssb    <- Fssb.m;
newFL@range  <- c(0, 1960, 0, 2004);

model(newFL) <- ricker();
```

<!---

The data below are derived from the [FLR tutorial](http://www.flr-project.org/doc/Loading_your_data_into_FLR.html), and the code below downloads the example data files into a temporary folder.

```{r}
#temp_dir <- tempdir()
#download.file("http://flr-project.org/doc/src/loading_data.zip", file.path(dir, "loading_data.zip"))
#unzip(file.path(temp_dir, "loading_data.zip"), exdir = temp_dir);
```

We can read in the example file below and convert it into a matrix to be used in an [FLQuant](http://www.flr-project.org/FLCore/reference/FLQuant.html) object. Rows in this matrix indicate ages.

```{r}
#fish_data <- read.csv(file.path(temp_dir, "catch_numbers.csv"), row = 1);
#fish_data <- as.matrix(fish_data);
#print(fish_data[,1:8]);
```

For simplicity, we merge these ages down into a single vector.

```{r}
fish_data <- apply(X = fish_data, MARGIN = 2, FUN = sum);
```

We can now create a new [FLQuant](http://www.flr-project.org/FLCore/reference/FLQuant.html) object with the `fish_data` vector. This process is important; down below, it will be repeated in a function that gets used within the GMSE population model, as GMSE will need to incorporate newly simulated data in each time step.

```{r}
fish_data_flq <- FLQuant(fish_data, dimnames = list(year = 1957:2011));
```

The data below include sample results for recruitment and spawning-stock biomass (SSB) of North Sea herring.

```{r}
data(nsher);
```

These data are structured using `FLStock`, an [S4](https://stat.ethz.ch/R-manual/R-devel/library/methods/html/Classes_Details.html) class of R objects. The code below creates an `FLSR` object, needed to create a stock-recruitment model.

```{r}
summary(nsher);
```

Notice that under `Model`, a Ricker model of stock-recruitment has already been fitted (many stock-recruitment models are available in FLCore; see the [FLSR tutorial](http://www.flr-project.org/doc/Modelling_stock_recruitment_with_FLSR.html) for more about fitting different models) and parameterised with Maximum Likelihood Estimation (MLE).  The code below shows how to apply the Beverton-Holt model model as an alternative, and the  `fmle` function parameterises the Beverton-Holt model model using the data in `nsher` and MLE.

```{r}
model(nsher) <- bevholt();
nsher        <- fmle(nsher);
summary(nsher);
```

We will use these tools to update the population model in GMSE.

a <- params(nsher)[[1]];
b <- params(nsher)[[2]];

rec(p4sr2);
ssb(p4sr2);

rec_year    <- as.numeric(attributes(rec(p4sr2))$dimnames$year);
ssb_year    <- as.numeric(attributes(ssb(p4sr2))$dimnames$year);
recruitment <- NULL;
for(year in 1:length(rec_year)){
    recruitment[year] <- rec(p4sr)[[year]];
}
spawning_stock_biomass <- NULL;
for(year in 1:length(ssb_year)){
    spawning_stock_biomass[year] <- ssb(p4sr)[[year]];
}

rec <- a * spawning_stock_biomass * exp(-b * spawning_stock_biomass);

rec(p4sr)[[51]] <- 4300

--->

>>>>

---
title: 'GMSE: an R package for generalised management strategy evaluation'
author: A. Bradley Duthie&#xb9;&#xb3;, Jeremy J. Cusack&#xb9;, Isabel L. Jones&#xb9;,
  Jeroen Minderman&#xb9;, Erlend B. Nilsen&#xb2;, Roc&#0237;o A. Pozo&#xb9;, 
  O. Sarobidy Rakotonarivo&#xb9;, Bram Van Moorter&#xb2;, and Nils Bunnefeld&#xb9;
date: '[1] Biological and Environmental Sciences, University of Stirling, Stirling,
  UK [2] Norwegian Institute for Nature Research, Trondheim, Norway [3] alexander.duthie@stir.ac.uk'
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    reference_docx: docx_template.docx
header-includes: null
linestretch: 1
link-citations: yes
linkcolor: blue
bibliography: gmse.bib
subtitle: Supporting Information 5
biblio-style: apalike
---

@@

Here we closely follow [a tutorial from the FLR project](http://www.flr-project.org/doc/Modelling_stock_recruitment_with_FLSR.html). To build the stock-recruitment model, the `FLCore` package is needed [@Kell2007].

```{r, eval = FALSE}
install.packages(c("FLCore"), repos="http://flr-project.org/R");
```

To start, we need to read in the `FLCore` library.

```{r}
library(FLCore);
```

For a simplified example in GMSE, we will simulate the process of stock recruitment over multiple time steps using an example stock-recruitment model. The stock-recruitment model describes the relationship between stock-recruitment and spawning stock biomass. The sample that we will work from is a recreation of the North Sea Herring (`nsher`) dataset available in the `FLCore` package [@Kell2007]. This data set includes recruitment and spawning stock biomass data between 1960 and 2004. First, we initialise an empty `FLSR` object and read in the recreated CSV data files.

```{r}
newFL        <- FLSR(); # Initialises the empty FLSR object
rec.n        <- read.csv("data/SI5/nsher_rec.csv");
ssb.n        <- read.csv("data/SI5/nsher_ssb.csv");
```

@@

We can then construct two `FLQuant` objects, specifying the relevant years and units.

```{r}
Frec.m       <- FLQuant(rec.m, dimnames=list(age=1, year = 1960:2004));
Fssb.m       <- FLQuant(ssb.m, dimnames=list(age=1, year = 1960:2004));
Frec.m@units <- "10^3";
Fssb.m@units <- "t*10^3";
```

We then place the recruitment and spawning stock biomass data into the `FLSR` object that we created.

```{r}
rec(newFL)    <- Frec.m;
ssb(newFL)    <- Fssb.m;
range(newFL)  <- c(0, 1960, 0, 2004);
```

The `FLCore` package offers several stock-recruitment models. Here we use a Ricker model of stock recruitment [@Ricker1954], and insert this model into the `FLSR` object below.

```{r}
model(newFL) <- ricker();
```

Parameters for the Ricker stock-recruitment model can be estimated with maximum likelihood.

```{r, echo = FALSE, messages = FALSE, cache = FALSE, comment = FALSE, results = "hide"}
newFL <- fmle(newFL);
```

```{r, eval = FALSE}
newFL <- fmle(newFL);
```

Diagnostic plots, identical to those of the [modelling stock-recruitment tutorial](http://www.flr-project.org/doc/Modelling_stock_recruitment_with_FLSR.html) for the `nsher_ri` example, are shown below.

```{r, fig.height = 7}
plot(newFL);
```

We now have a working example of a stock-recruitment model, but for our integration with `gmse_apply`, we will want a function that automates the above to simulate the process of updating the stock-recruitment model. We do this using the custom function created below.

```{r}
update_SR_model <- function(rec_m, ssb_m, years){
    Frec_m       <- FLQuant(rec.m, dimnames=list(age=1, year = years));
    Fssb_m       <- FLQuant(ssb.m, dimnames=list(age=1, year = years));
    Frec_m@units <- "10^3";
    Fssb_m@units <- "t*10^3";
    rec(newFL)   <- Frec.m;
    ssb(newFL)   <- Fssb.m;
    range(newFL) <- c(0, years[1], 0, years[length(years)]);
    model(newFL) <- ricker();
    newFL        <- fmle(newFL);
    return(newFL);
}
```

The above function will be used within another custom function to predict the next time step of recruitment.

```{r}
predict_recruitment <- function(rec_m, ssb_m, years, new_ssb){
    newFL <- update_SR_model(rec_m, ssb_m, years);
    a     <- params(newFL)[[1]] # Extract 'a' parameter of the Ricker model
    b     <- params(newFL)[[2]] # Extract 'b' parameter of the Ricker model
    rec   <- a * new_ssb * exp(-b * new_ssb); # Predict the new recruitment
    return(recruitment)
}
```

In `gmse_apply`, we will use the `predict_recruitment` function above as the resource (i.e., operational) model. The `new_ssb` reads in the new spawning stock biomass, which will be calculated from the built-in GMSE `user` model.

Integrating `predict_recruitment` with `gmse_apply`
================================================================================
